\Huge Conferences in an Era of Expensive Carbon

### 2019-06-17 main.tex
CP
New outline

\begin{itemize}
\item Accountability.
  \begin{itemize}
  \item Proposal: Every ACM-sponsored conference should be required to
  publicly report its carbon footprint (total and per-participant).
  (Including PC meetings, participant travel, maybe hotel?).
  \item ACM needs to provide tools to help conferences do this.
  \item Benefits:
  \begin{itemize}
  \item Encourages people to pay attention, which will lead to some changes
  in behavior all by itself. 
  \item Generates data that will be needed for grounding discussions of how
  optimize footprint.
  \item Indeed, optimizing footprint is not so easy.  (Describe some of our
  data.)  More data and thorough analysis is needed.
  \end{itemize}
  \end{itemize}
\item Pricing carbon.
  \begin{itemize}
  \item To motivate more painful changes and force people to make decisions,
  we need a price on carbon.
  \item Ultimately, we want/expect governments to do this.  But we don't need to
  wait. 
  \item Proposal: ACM should apply downward pressure on conferences by
  charging conferences a surcharge based on their carbon footprint.
  \item This money can be used in many ways: buying carbon offsets,
  supporting ``green research'', supporting virtualizing conferences
  (livestreaming, etc.), supporting a transition to a low-carbon / few
  conferences business model for ACM.
  \end{itemize}
\end{itemize}

A broad scientific consensus warns that human
emissions of carbon dioxide and other greenhouse gases are warming the earth.
% (Figure \ref{fig:warming}).
This is not a problem we can leave to future
generations: The UN's Intergovernmental
Panel on Climate Change (IPCC) says a 40\% decrease in
emissions by 2030 is needed to avoid irreversible damage~\cite{IPCC-report}.
% (Figure~\ref{fig:IPCC-cover}).
Achieving reductions on this scale will
require urgent and sustained commitment at all levels of society, including
not only city, state, and national governments, but also 
universities, companies, and scientific societies
like ACM.

@@

Indeed, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger~\cite{carboncalculator}\iflater\bcp{Trim citation?}\fi---a
significant fraction of the {\em total} yearly emissions for an average
resident of the US (16.5 tons) or Europe (7 tons)~\cite{emissions}.
\iflater\bcp{For comparison, ...  (put in some figures about the footprints
  of a few other things that people might think of reducing---driving,
  eating meat, upgrading their fridge...)}\fi{} Moreover, these emissions
have no near-term technological fix, since jet fuel is difficult to replace
with renewable energy sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

On the other hand, conferences obviously also create significant value.
Science is a fundamentally social process, and the conference system
accelerates research through high-bandwidth interaction, direct
dissemination of results, network building, and serendipitous
cross-fertilization.

This presents a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  It seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.

@@

Fortunately, there is a well-established mechanism for comparing apples and
oranges---indeed, for comparing almost anything!  Economists call it ``the
magic of markets.''  By mapping choices and preferences in disparate domains
onto a common scale (i.e., money), economic markets function as massively
distributed computations that optimize costs and benefits across vastly
different activities.\bcp{Give a simple, concrete example.}  For this
reason, many policy experts advocate addressing climate change through some
form of carbon pricing, which places a monetary cost on greenhouse gas
emissions, sending a clear market signal that changes aggregate
behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}  Exactly how the price of carbon
should be set is a challenging policy question, but it has a clear goal: It
should be chosen so that the world arrives at zero emissions within the
timeframe that climate scientists tell us is safe.

@@

>>>>

A broad scientific consensus warns that human
emissions of carbon dioxide and other greenhouse gases are warming the earth.
% (Figure \ref{fig:warming}).
This is not a problem that can be left to future
generations: The UN's Intergovernmental
Panel on Climate Change (IPCC) says a 40\% decrease in
emissions is needed by 2030 to avoid irreversible damage~\cite{IPCC-report}.
% (Figure~\ref{fig:IPCC-cover}).
Achieving reductions on this scale will
require urgent and sustained commitment at all levels of society, including
not only city, state, and national governments, but also 
universities, companies, and scientific societies
like ACM.

@@

In fact, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger~\cite{carboncalculator}\iflater\bcp{Trim citation?}\fi---a
significant fraction of the {\em total} yearly emissions for an average
resident of the US (16.5 tons) or Europe (7 tons)~\cite{emissions}.
\iflater\bcp{For comparison, ...  (put in some figures about the footprints
  of a few other things that people might think of reducing---driving,
  eating meat, upgrading their fridge...)}\fi{} Moreover, these emissions
have no near-term technological fix, since jet fuel is difficult to replace
with renewable energy sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

How should ACM respond to these facts?  We have two concrete proposals.

\SECTION{Mandate public accounting of carbon footprints.}  
\bcp{The section headers could use some wordsmithing!}

  \begin{itemize}
  \item Proposal: Every ACM-sponsored conference should be required to
  publicly report its carbon footprint (total and per-participant).
  (Including PC meetings, participant travel, maybe hotel?).
  \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
  built one.)
  \item Benefits:
  \begin{itemize}
  \item Encourages people to pay attention, which will lead to some changes
  in behavior all by itself. 
  \item Generates data that will be needed for grounding discussions of how
  optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
  meetings.  Livestreaming.
  \end{itemize}
\end{itemize}

These are things that we can and should be doing now.  Indeed, many parts
of ACM are already doing them.  (In SIGPLAN, we have... and the experience
has been...)

However, going beyond these simple steps is going to be much more
challenging.  One reason is that it isn't actually very clear what changes
will help significantly.  (Describe some of our data.)  More data and
thorough analysis is needed.

Another reason why going beyond simple measures is hard is that, although
conferences damage the climate, they also create significant value.  Science
is a fundamentally social process, and the conference system accelerates
research through high-bandwidth interaction, direct dissemination of
results, network building, and serendipitous cross-fertilization.

This presents a puzzle: How can we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  The question is difficult because the harm and
the compensating benefits cannot be directly compared---they are apples and
oranges.

Fortunately, there is a well-established mechanism for comparing apples and
oranges---indeed, for comparing almost anything!  Economists call it ``the
magic of markets.''  By mapping choices and preferences in disparate domains
onto a common scale (i.e., money), economic markets function as massively
distributed computations that optimize costs and benefits across vastly
different activities.\bcp{Give a simple, concrete example.}  For this
reason, many policy experts advocate addressing climate change through some
form of carbon pricing, which places a monetary cost on greenhouse gas
emissions, sending a clear market signal that changes aggregate
behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}

\SECTION{Put a Price on Carbon.}

  \begin{itemize}
  \item To motivate more painful changes and force people to make difficult
  choices, we need a price on carbon.
  \item Ultimately, governments should do this.  But we don't need to wait:
  as several companies have shown, even an artificial ``internal price'' on
  carbon creates a significant price signal that shapes behavior.
  \item Our second (more ambitious) proposal is: ACM should begin requiring
  conferences to pay a surcharge based on their carbon footprint.  It should
  be low at the beginning and increase steadily (and predictably) year on year.
  \item Benefit: This will create concrete downward pressure on conferences
  to decrease their footprints.  They can then choose (on the basis of
  careful analysis of data and weighing of benefits) how best to do so.
  \item The money raised in this way can be used in many ways: buying carbon
  offsets, supporting ``green computing research'', virtualizing conferences
  (livestreaming, etc.), supporting a transition to a low-carbon / few
  conferences business model for ACM.
  \end{itemize}

(Some kind of closing.)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\section*{Old text (that may be useful to mine)}

@@

Exactly how the price of carbon
should be set is a challenging policy question, but it has a clear goal: It
should be chosen so that the world arrives at zero emissions within the
timeframe that climate scientists tell us is safe.

@@

\bcp{New outline:
\begin{itemize}
\item Accountability.
  \begin{itemize}
  \item Proposal: Every ACM-sponsored conference should be required to
  publicly report its carbon footprint (total and per-participant).
  (Including PC meetings, participant travel, maybe hotel?).
  \item ACM needs to provide tools to help conferences do this.  (SIGPLAN
  built one.)
  \item Benefits:
  \begin{itemize}
  \item Encourages people to pay attention, which will lead to some changes
  in behavior all by itself. 
  \item Generates data that will be needed for grounding discussions of how
  optimize footprint.  E.g., SIGPLAN has mostly gone to virtual PC
  meetings.  Livestreaming.
  \item Indeed, optimizing footprint is not so easy.  (Describe some of our
  data.)  More data and thorough analysis is needed.
  \item Segue: Colocation, etc.
  \end{itemize}
  \end{itemize}
\item Pricing carbon.
  \begin{itemize}
  \item To motivate more painful changes and force people to make decisions,
  we need a price on carbon.
  \item Ultimately, we want/expect governments to do this.  But we don't need to
  wait. 
  \item Proposal: ACM should apply downward pressure on conferences by
  charging conferences a surcharge based on their carbon footprint.
  \item This money can be used in many ways: buying carbon offsets,
  supporting ``green research'', supporting virtualizing conferences
  (livestreaming, etc.), supporting a transition to a low-carbon / few
  conferences business model for ACM.
  \end{itemize}
\end{itemize}
}
